---
title: "607_AssignmentW5"
author: "Ariann Chai"
date: "2023-09-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data 607 - Assignment W5

For this assignment, we had to import some flight data and change the dataframe into a tidy one (using the tidyr library). So first, I started by getting my libraries. 

```{r libraries}
library(dplyr)
library(tidyr)
library(readr)
```

##Importing the data
For my R file and work, I used mySqlConnection as shown below. 

```{sqlconnection}
install.packages("RMySQL")
library(RMySQL)

mysqlconnection = dbConnect(RMySQL::MySQL(),
                            dbname='607_AssignmentW5',
                            host='localhost',
                            port=3306,
                            user='root',
                            password='')
```

However, since I had to delete the password for this command and it will not work, I saved it as a csv to be called for this RMarkdown file

```{arrivalDelays query}
result = dbSendQuery(mysqlconnection, "select * from arrivalDelays") 
arrivalDelays = fetch(result)
write.csv(arrivalDelays,file='/Users/Ari/Data607/assignmentW5/arrivalDelays.csv')
```

So now we can grab it as a csv. We also need to get rid of the extra id column that was generated from the sql table.
```{r arrivalDelays csv}
arrivalDelays <- read_csv("arrivalDelays.csv")

#clean up the data and set it as a simple dataframe
arrivalDelays <- subset(arrivalDelays, select = -1)
as_tibble(arrivalDelays)
```

## Cleaning the data
Now we must change this dataframe into a tidy one. From the lecture and readings, I learned a tidy table means "1 row for each observations".  

Right now, the above table contains 2 airlines that have 2 rows each for 2 different delay status (on time, delayed) but, each row contains 5 different observations in the format of the 5 columns based on locations (LosAngeles, Phoenix, SanDiego, SanFrancisco, Seatte). We need to separate each row into 5 rows that reflection each of these locations. In the end, we will have 4 (2 airlines * 2 delay status) * 5 locations = 20 rows.  

In order to do this, I used the pivot_longer command and had it separate the rows based on the 5 location columns. They are then combined into 2 new columns: 1 for the location name and 1 for the value (# of flights) that used to sit in the individual location column.
```{r arrivalDelay_tidy}
arrivalDelay_tidy <- 
  pivot_longer(arrivalDelays, 
             cols=c('LosAngeles', 'Phoenix', 'SanDiego', 'SanFrancisco', 'Seatte'),
             names_to = 'Locations',
             values_to = 'Flights')
arrivalDelay_tidy <- arrivalDelay_tidy[order(arrivalDelay_tidy$Locations),]
arrivalDelay_tidy
```

## Anaylsis

```{r arrivalDelay_tidy_airline}
arrivalDelay_tidy %>% group_by(airline) %>% 
  summarise(count=sum(Flights),
            .groups = 'drop')

arrivalDelay_tidy %>% 
  group_by(airline, delayStatus) %>% 
  summarise(count=sum(Flights),
            .groups = 'drop')
```

The two tables above show us the number of flights from each airline and the breakdown of that airlines' on-time & delayed flights. AM WEST has more flights than ALASKA and less delayed flights as when you work out the ratios: 1 in 6.5 flights from ALASKA are delayed and 1 in 8.2 flights from AM WEST are delayed.

```{r arrivalDelay_tidy_location}
arrivalDelay_tidy %>% 
  group_by(Locations) %>% 
  summarise(count=sum(Flights),
            .groups = 'drop')

arrivalDelay_tidy %>% 
  group_by(Locations, delayStatus) %>% 
  summarise(count=sum(Flights),
            .groups = 'drop')
```

The two tables above show us the number of flights arriving to each location and the breakdown of that location's on-time & delayed arrived flights. Phoenix has the most arrival flights and least amount of delayed flights is San Francisco. The ratio breakdown of delayed flights to on-time flights is below:  
- LosAngeles: 1 in 6.6   
- Phoenix: 1 in 11.8  
- SanDiego: 1 in 7  
- SanFrancisco: 1 in 3.6  
- Seattle: 1 in 5.6  

