---
title: "607 Project 1"
author: "Ariann Chai"
date: "2023-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is the R Markdown document for my 607 project 1. I had to take in txt file (tournamentinfo.txt) of chess tournament data and make a new csv with 5 columns using the info in the txt file:

- Player’s Name  
- Player’s State  
- Total Number of Points  
- Player’s Pre-Rating  
- Average Pre Chess Rating of Opponents  

In order to accomplish this, I saved the 5 requested columns into separate vectors and combined them into a dataframe later. To get the needed data for the vectors, I heavily used str_extract and regex/pattern making.  

Within the used regexs, the way the data was saved (with | in between columns) was a big help in targeting specific sections of the string. Working with str_extract gave a lot of NA values as well that needed to get removed with !is.na().

```{r libraries}
library(stringr)
library(readr)
```

## Load txt file
First step was to load in the txt file, tournamentinfo.txt with the needed data. Below is what the raw data from that file looked like:
```{r load_txt}
txt_data <- paste(readLines("tournamentinfo.txt"))
txt_data
```

## Make and fill a vector for Player's Name
I used str_extract with the regular expression "\\| [A-Z]+(.)+[A-Z]+     ". The (.) and additional spaces after were added in order to catch full names of people with more than 2 words i.e. "THOMAS JOSEPH HOSMER".  
This is the only vector that had an extra row in it as the label at the top of the txt file with "USCF ID" was also returning in player_names so it had to be manually removed.
```{r player_names}
player_names <- str_extract(txt_data, "\\| [A-Z]+(.)+[A-Z]+     ")
player_names <-player_names[!is.na(player_names)]
player_names <- str_extract(player_names, "[A-Z]+(.)+[A-Z]+")
player_names <- player_names[ !player_names == 'USCF ID']
player_names
```

## Make and fill a vector for Player's State
I used str_extract with the regular expression "[A-Z][A-Z] \\|". 
This one was simplier than the player's names as states are always two 2 letters.
```{r player_states}
player_states <- str_extract(txt_data, "[A-Z][A-Z] \\|")
player_states <-player_states[!is.na(player_states)]
player_states <- str_extract(player_states, "[A-Z][A-Z]")
player_states
```

## Make and fill a vector for Total Number of Points
I used str_extract with the regular expression "\\|[0-9].[0-9]"
This one was simple like player's state as it would always be #.#. 
I could also write the regex as "\\|[0-9].[0,5]" as the 2nd number was always 0 or 5.
```{r points_total}
points_total <- str_extract(txt_data, "\\|[0-9].[0-9]")
points_total <-points_total[!is.na(points_total)]
points_total <- str_extract(points_total, "[0-9].[0-9]")
points_total
```

## Make and fill a vector for Player’s Pre-Rating
I used str_extract with the regular expression "R: +[0-9][0-9][0-9]+".
```{r player_prerating}
player_prerating <- str_extract(txt_data, ("R: +[0-9][0-9][0-9]+"))
player_prerating <-player_prerating[!is.na(player_prerating)]
player_prerating <- str_extract(player_prerating, "[0-9][0-9][0-9]+")
player_prerating
```

## Make and fill a vector for Average Pre Chess Rating of Opponents
This last vector was a lot more work than the other vectors as we first had to extract the section of each player's data with their opponents' ids. It took 2 str_extracts to get the opponent ids by themselves (together in a character vector for that player). Below is the code with the str_extract statements and examples of what a line in opponents_prerating looks like.
```{r opponents_prerating}
opponents_prerating <- str_extract(txt_data, ("(\\|[L,W,D]+ +[0-9]+).*\\|"))
opponents_prerating <- opponents_prerating[!is.na(opponents_prerating)]
opponents_prerating[1]

opponents_prerating <- str_extract_all(opponents_prerating,"[0-9]+")
opponents_prerating[1]
```  
After getting it down to just the opponents' id numbers for each player, I then needed to find the mean of each player's opponents pre rating. In order to do this, I used 2 for loops (1 looping through the players and 1 looping around the opponents' id for that player). For each opponent id in that player's vector, I called the vector player_prerating I made earlier to get the opponent's prerating. After the last opponent id in that player's vector, I divided it by number of opponents and added it to avg_opponents_prerating. I made use of two temp variables for the number of opponents for each player and the combined prerating scores of their opponents that would reset after every player vector.
```{r avg_opponents_prerating}}
avg_opponents_prerating <- vector(length=length(opponents_prerating))
for (i in 1:length(opponents_prerating)){
  opponents_num <- 0 
  scores_combined <- 0 
  for (j in opponents_prerating[[i]]){
    opponents_num <- opponents_num+1
    scores_combined <- scores_combined+as.numeric(player_prerating[as.numeric(j[1])])
  }
  avg_opponents_prerating[i] <- scores_combined/opponents_num
}
avg_opponents_prerating <- round(avg_opponents_prerating, 0)
avg_opponents_prerating
```

## Turn the 5 vectors into a dataframe
With all 5 vectors created and filled, it is now time to make the csv_data dataframe that will be saved as a csv later. 
Below is the final results of the collected data from tournamentinfo.txt
```{r csv_data}
csv_data <- data.frame(player_names,player_states,points_total,player_prerating,avg_opponents_prerating)

colnames(csv_data)[1]="Player’s Name"
colnames(csv_data)[2]="Player’s State"
colnames(csv_data)[3]="Total Number of Points"
colnames(csv_data)[4]="Player’s Pre-Rating"
colnames(csv_data)[5]="Average Pre Chess Rating of Opponents"

csv_data
```

## Write dataframe to csv
For the final step, using the readr library and the function write_csv, we can write csv_data into tournamentinfo.csv.
```{r write_csv}
write_csv(csv_data, "tournamentinfo.csv")
```

